---
title: "Solution Exercise 7 - Dynamic lot sizing"
author: "Thomas Kirschstein"
subtitle: Inventory Management
output:
  pdf_document: default
  html_document:
    df_print: paged
    toc: yes
    toc_float: true
    number_sections: false
---



```{r, setup, echo=F, include=T, error=FALSE, warning=FALSE, message=FALSE}
library(DT)
library(kableExtra)
library(knitr)

# Normal loss functions - direct
v.fun.direct.norm <- function(s, mu = 0, sigma = 1){
  integrate(function(y) (y-s)*dnorm(y, mean=mu, sd = sigma), lower = s, upper=Inf)$value
}

# deviation Beta vs. exp. loss 
v.diff.norm <- function(x, beta, mu = 0, sigma = 1, L,  c.or, c.sh){
  (v.fun.direct.norm(x[1], mu = L*mu, sigma = sqrt(L)*sigma) - (1-beta)*x[2])
}


v.diff.norm.q <- function(s, q, beta, mu = 0, sigma = 1, L,  c.or, c.sh){
  (v.fun.direct.norm(s, mu = L*mu, sigma = sqrt(L)*sigma) - (1-beta)*q)^2
}



v.fun.direct.gamma <- function(s, mu = 0, sigma = 1){
  integrate(function(y) (y-s)*dgamma(y, shape = mu^2/sigma^2, rate = mu/sigma^2), lower = s, upper=Inf)$value
}

v.fun.gamma <- function(s, mu = 0, sigma = 1){
  alpha <- mu^2/sigma^2
  beta <- mu/sigma^2
  alpha/beta * (1 - pgamma(s, shape = alpha+1, rate = beta)) - s*(1 - pgamma(s, shape = alpha, rate = beta))
}

# beta-Servicelevel
v.fun <- function(x, mu = 0, sigma = 1) {
  sigma * (dnorm((x-mu)/sigma) - (x-mu)/sigma * (1-pnorm((x-mu)/sigma) ))
}
# optimizer beta SL
v.fun.opt.loss <- function(x, loss = loss.norm, ...) (loss - v.fun(x, ...))^2 


# shortest path / WW

ww.opt.func <- function(c.mat){
  n <- nrow(c.mat)
  res.mat <- matrix(NA, ncol=n, nrow=n)
  res.id <- numeric(n)
  # initializing
  res.mat[1,] <- c.mat[1,]
  for(i in 2:n){
    res.id[i-1] <- which.min(res.mat[,i-1])
    res.mat[i,] <- c.mat[i,] + res.mat[res.id[i-1],i-1]
  }
  res.id[n] <- which.min(res.mat[,n])
  polic <- res.id[n]
  repeat{
    polic <- c(polic , res.id[tail(polic,1)-1] )
    if(tail(polic,1) == 1) break
  }
  polic <- rev(polic)
  
  return(list(cum.cost = res.mat, ord.per = polic, ids = res.id))
}



```

# Dynamic lotsizing with $\alpha$ service level constraint

1. Calculate the matrices of means, standard deviations and  order-up levels $S$. Derive the total cost matrix. 

```{r, exe-1-1, echo=F, message=F, error=FALSE , include=T}
options(knitr.kable.NA = '')
# solution
t.vec <- 1:6
mu.vec <- round(60 + 10*sin(2*pi*t.vec/max(t.vec)) + 30*cos(2*pi*t.vec/max(t.vec)),0)
sig.vec <- mu.vec * 0.2

c.or <- 10
c.sh <- 0.022
alpha <- .99 #0.9009009
n <- length(mu.vec)

seqs <- n:1
# means
mean.mat <- matrix(NA, ncol=n, nrow=n, byrow=T) 
mean.mat[lower.tri(mean.mat, diag = T)] <- unlist(sapply(seqs, function(x) cumsum(tail(mu.vec, x) ) ))
mean.mat <- t(mean.mat)

df <- as.data.frame(round(mean.mat,2))
colnames(df) <- rownames(df) <- 1:n
kable(df, "pandoc", caption = "Mean matrix", row.names = T)

# sds
sds.mat <- matrix(NA, ncol=n, nrow=n, byrow=T) 
sds.mat[lower.tri(mean.mat, diag = T)] <- unlist(sapply(seqs, function(x) sqrt(cumsum(tail(sig.vec^2, x) ) ) ))
sds.mat <- t(sds.mat)

df <- as.data.frame(round(sds.mat,2))
colnames(df) <- rownames(df) <- 1:n
kable(df, "pandoc", caption = "Standard deviation matrix", row.names = T)

z <- qnorm(alpha)

# S mat
S.mat <- mean.mat + z*sds.mat

df <- as.data.frame(round(S.mat,2))
colnames(df) <- rownames(df) <- 1:n
kable(df, "pandoc", caption = "Order-up level (S) matrix", row.names = T)

# cost mat
mult.mat <- matrix(NA, ncol=n, nrow=n) 
mult.mat[lower.tri(mult.mat, diag = T)] <- unlist( sapply(1:n, function(x) tail(1:n,n-x+1)-x+1 ) )
mult.mat <- t(mult.mat)

mmeans.mat <- matrix(NA, ncol=n, nrow=n) 
mmeans.mat[lower.tri(mmeans.mat, diag = T)] <- unlist(sapply(seqs, function(x) sapply(1:x , function(y) sum(y:1 * head(tail(mu.vec, x),y) ) )))
mmeans.mat <- t(mmeans.mat)

cost.mat <- c.or + c.sh * (mult.mat * S.mat -  mmeans.mat)

df <- as.data.frame(round(cost.mat,2))
colnames(df) <- rownames(df) <- 1:n
kable(df, "pandoc", caption = "Partial cost matrix", row.names = T)

```

2. Determine the optimal weekly replenishment strategy and calculate expected stock levels as well as order quantities. On which weekdays has retailer to expect potential shortfalls?

Based on the partial cost, the cumulative total cost are as follows:
```{r, exe-1-2, echo=F, message=F, error=FALSE , include=T}
res.1 <- ww.opt.func(cost.mat)
df <- as.data.frame(round(res.1$cum.cost,2))
colnames(df) <- rownames(df) <- 1:n
kable(df, "pandoc", caption = "Total cost matrix", row.names = T)
```

Thus, the optimal policy is to order in periods `r paste(res.1$ord.per, collapse = " and ")`. Thus, order-up levels are $S_1 =$ `r round(S.mat[1,4],2)` and $S_5=$ `r round(S.mat[5,6],2)`. For expected inventory levels and expected order quantities follows:

```{r, echo=FALSE}
i.vec <- c(round(S.mat[1,4],2) - mean.mat[1,1:4], round(S.mat[5,6],2) - mean.mat[5,5:6])
q.vec <- c(round(S.mat[1,4],2), 0,0,0, round(S.mat[5,6],2)-i.vec[4], 0)
df <- as.data.frame(rbind(q.vec,i.vec))
colnames(df) <- 1:n
rownames(df) <- c("$q_t$","$i_t$")
kable(df, "pandoc", row.names = T)
```

Shortfalls may (most likely) occur in period 4 and 6 (on Thursday and Saturday).

3. A consultant suggests that the retail manager should incorporate shortage cost instead of the $\alpha$ service level to determine the optimal ordering policy. The retail manager estimates a shortage cost rate of 0.20 Euro per unit of milk. What happens to the replenishment solution? (no calculation required)

With the given  cost rates a critical ratio of $CR=\frac{c^{sf}}{c^{sf} + c^{sh}}=\frac{0.20}{0.20 + 0.022} \approx 90.1$%. Thus, the safety multiplier $z$ reduces to `r round(qnorm(0.901),2)`. Consequently, the order-up levels $S$ decrease linearly w.r.t. $\sigma$ which also holds for the partial cost. As the ratio between means and standard deviations is comparatively small ($c=0.2$), the structure of partial costs is probably not much affected. Therefore, the solution probably remains the same (same ordering periods), but the total cost decrease. 

# Dynamic lotsizing with $\beta$ service level constraint

The following table displays expected demand and standard deviation of a material for the next 7 days. 

| $t$        | 1   | 2  |  3 |   4  |  5  |  6  |  7  |
|:------:|:------:|:------:|:------:|:------:|:------:|:------:|:------:|
| $\mu_t$     | 120 | 75 | 82 | 100 | 91 | 65 | 88 |
| $\sigma_t$  | 30 | 12 | 25 | 10 | 28 | 5 | 10 |

The demand is assumed to be independently normally distributed. The manager intends to assure a $\beta$ service level of 98%. Ordering cost are $c^{or}=150$ Euro and stock-holding cost rate $c^{sh} = 0.5$ Euro per unit and period.

1. Calculate the matrices of means, standard deviations and  order-up levels $S$. Derive the total cost matrix. 

```{r, exe-2-1, echo=F, message=F, error=FALSE , include=T}
# solution
mu.vec <- c(120,75,82,100,91,65,88)
sig.vec <- c(30,12,25,10,28,5,10)

c.or <- 150
c.sh <- 0.5
beta <- .97 #0.9009009
n <- length(mu.vec)

seqs <- n:1
# means
mean.mat <- matrix(NA, ncol=n, nrow=n, byrow=T) 
mean.mat[lower.tri(mean.mat, diag = T)] <- unlist(sapply(seqs, function(x) cumsum(tail(mu.vec, x) ) ))
mean.mat <- t(mean.mat)
# sds
sds.mat <- matrix(NA, ncol=n, nrow=n, byrow=T) 
sds.mat[lower.tri(mean.mat, diag = T)] <- unlist(sapply(seqs, function(x) sqrt(cumsum(tail(sig.vec^2, x) ) ) ))
sds.mat <- t(sds.mat)

# S mat
S.mat <- matrix(NA, ncol=n, nrow=n) 
S.mat[upper.tri(S.mat, diag = T)] <- 
  sapply(which(upper.tri(mean.mat, diag = T)), function(x) {
  optim(v.fun.opt.loss, loss = mean.mat[x]*(1-beta), mu = mean.mat[x], sigma = sds.mat[x], par = mean.mat[x], lower = 0, upper = mean.mat[x]+5*sds.mat[x], method="L-BFGS-B")$par
})


# cost mat
mult.mat <- matrix(NA, ncol=n, nrow=n) 
mult.mat[lower.tri(mult.mat, diag = T)] <- unlist( sapply(1:n, function(x) tail(1:n,n-x+1)-x+1 ) )
mult.mat <- t(mult.mat)

mmeans.mat <- matrix(NA, ncol=n, nrow=n) 
mmeans.mat[lower.tri(mmeans.mat, diag = T)] <- unlist(sapply(seqs, function(x) sapply(1:x , function(y) sum(y:1 * head(tail(mu.vec, x),y) ) )))
mmeans.mat <- t(mmeans.mat)

cost.mat <- c.or + c.sh * (mult.mat * S.mat -  mmeans.mat)
#
res.2 <- ww.opt.func(cost.mat)

df.mean <- as.data.frame(round(mean.mat,2))
colnames(df.mean) <- rownames(df.mean) <- 1:n
kable(df.mean, "pandoc", caption = "Mean matrix", row.names = T)

df.sd <- as.data.frame(round(sds.mat,2))
colnames(df.sd) <- rownames(df.sd) <- 1:n
kable(df.sd, "pandoc", caption = "Standard deviation matrix", row.names = T)

df.S <- as.data.frame(round(S.mat,2))
colnames(df.S) <- rownames(df.S) <- 1:n
kable(df.S, "pandoc", caption = "Order-up level (S) matrix", row.names = T)

df.cost <- as.data.frame(round(cost.mat,2))
colnames(df.cost) <- rownames(df.cost) <- 1:n
kable(df.cost, "pandoc", caption = "Partial cost matrix", row.names = T)
```


2. Determine the optimal weekly replenishment strategy and calculate expected stock levels as well as order quantities. 

Based on the partial cost, the cumulative total cost are as follows:
```{r, exe-2-2, echo=F, message=F, error=FALSE , include=T}
res.2 <- ww.opt.func(cost.mat)
df <- as.data.frame(round(res.2$cum.cost,2))
colnames(df) <- rownames(df) <- 1:n
kable(df, "pandoc", caption = "Total cost matrix", row.names = T)
```

Thus, the optimal policy is to order in periods `r paste(res.2$ord.per, collapse = " and ")`. Thus, order-up levels are $S_1 =$ `r round(S.mat[1,3],2)` and $S_4=$ `r round(S.mat[4,7],2)`. For expected inventory levels and expected order quantities follows:

```{r, echo=FALSE}
i.vec <- c(round(S.mat[1,3],2) - mean.mat[1,1:3], round(S.mat[4,7],2) - mean.mat[4,4:7])
q.vec <- c(round(S.mat[1,3],2), 0,0, round(S.mat[4,7],2)-i.vec[3], 0,0,0)
df <- as.data.frame(rbind(q.vec,i.vec))
colnames(df) <- 1:n
rownames(df) <- c("$q_t$","$i_t$")
kable(df, "pandoc", row.names = T)
```


3. Assume the supplier can ship only 200 units at most. Does the optimal solution change? Try to find an alternative solution by adapting the Wagner-Whithin algorithm.

Yes, the solution needs to be changed as the expected order quantities are larger than the transport capacity. Alternative solution with at least 3 orders?

