---
title: "Solutions Exercise 4 - Newsvendor"
author: "Thomas Kirschstein"
subtitle: Inventory Management
output:
  pdf_document: default
  html_document:
    df_print: paged
    toc: yes
    toc_float: true
    number_sections: false
---
```{r, setup, echo=F, include=T}
library(DT)
library(kableExtra)
library(knitr)
```

# Newsvendor (I)

Data: $c_u = 1.95 - 0.80 = 1.15$ Euro and $c_o = 0.80 - 0.10 = 0.70$ Euro with $X \sim N(22, 6^2)$

1. critical ratio $CR = \frac{c_u}{c_u + c_o} = \frac{1.15}{1.15 + 0.70}=0.6216216$ $\Rightarrow$ $q^* = 22 + 6 \cdot 0.30974 = 23.858 \approx 24$ loafs is optimal

2. expected cost: $E(Z) = (c_u + c_o) \cdot \varphi(z^*) \cdot \sigma = 1.85 \cdot 6 \cdot \varphi(0.30974) =  4.22$ Euro
expected profit: $G(Z) = c_u \cdot \mu - (c_u + c_o) \cdot \varphi(z^*) \cdot \sigma = 1.15 \cdot 22 - 1.85 \cdot 6 \cdot \varphi(0.30974) =  21.07$ Euro

3. $\alpha$ SL equals the critical ratio, i.e. 62%.
$\beta$ SL is $1-\frac{L(Y,q)}{\mu}=1-\sigma \cdot \frac{L(Z,q')}{\mu}$. Thus, with $q' = 0.30974$ and $\sigma = 6$ it follows  $\beta = 1-6 \cdot \frac{0.26}{22} = 92.9$%.

# Newsvendor (II)
\begin{align*}
Z (q)=&  c_o \cdot \int_{0}^{q} (q-y) \cdot f_y(y) dy + c_u \cdot \int_{q}^{\infty} (y-q) \cdot f_y(y) dy & \text{where } y \sim U\left(75, 125\right)\\
=&  c_o \cdot \int_{75}^{q} (q-y) \cdot 0.02 dy + c_u \cdot \int_{q}^{125} (y-q) \cdot 0.02 dy & \\
=&  c_o \cdot \left[ 0.02 \cdot q \cdot y - 0.01 \cdot y^2 \right]_{75}^{q}  + c_u \cdot \left[  0.01 \cdot y^2 - 0.02 \cdot q \cdot y \right]_{q}^{125} &   \\
=&  0.3 \cdot q^2 - 45 \cdot q +1687.5 + 9375 - 150 \cdot q + 0.6\cdot q^2 & \text{with }c_u = 60 \text{ and }c_o=30\\
=&  0.9 \cdot q^2 - 195 \cdot q + 11062.5&
\end{align*}

The derivative of $Z(q)$ is $Z'(q) = 1.8 \cdot q - 195$ such that $q^* \approx 108$. The expected profit is $100 \cdot 60 - Z(108) = 600 - 500.1 \approx 100$ Euro.

# Discrete newsvendor

The spare part demand $Y$ for solar aggregates is binomially distributed with $n=20 \cdot 5 = 100$ and $p=0.04$, i.e. $Y \sim B(n,p)$. Thus, the density values can be calculated as $P(Y=y) = {n \choose y} \cdot p^y \cdot (1-p)^{n-y}$.

Total cost are calculated by $E(C(q))=c^u \cdot (\mu - q) + (c^u + c^o ) \cdot E(S^+(q))$. Undershooting cost and overshooting cost are calculated as $E(C^{us}(q)) = c^u \cdot  E(S^-(q)) = c^u \cdot \sum_{y=q}^{15} (y-q)\cdot f(y)$ and  $E(C^{os}(q)) = c^o \cdot  E(S^+(q)) = c^u \cdot \sum_{y=0}^{q} (y-q)\cdot f(y)$.

```{r, exe4-3, echo=FALSE}
n <- 100
p <- 0.04
cu <- 5
co <- 1
x.vals <- 0:15

res.mat <- data.frame("y/q" = x.vals, 
                      density = round(dbinom(x.vals, size = n, prob = p),4),
                      cost.os = NA, cost.us = NA, total.cost = NA)
for(i in 1:nrow(res.mat)){
  tmp.p <- res.mat$y[i]
  tmp.os <- sum((tmp.p - res.mat$y[1:(i-1)])*res.mat$dens[1:(i-1)])
  tmp.us <- sum((res.mat$y[i:nrow(res.mat)] - tmp.p)*res.mat$dens[i:nrow(res.mat)])
  res.mat$cost.os[i] <- co * tmp.os
  res.mat$cost.us[i] <- cu * tmp.us
  res.mat$total.cost[i] <- tmp.os*(cu+co) + cu*(n*p - tmp.p  )  #tmp.us*cu + tmp.os*co
}
# checking cost calculation approaches
#(res.mat$cost.os+res.mat$cost.us) - res.mat$total.cost 
res.mat$cost.os <- round(res.mat$cost.os,2) 
res.mat$cost.us <- round(res.mat$cost.us,2) 
res.mat$total.cost <- round(res.mat$total.cost,2) 
kable(res.mat , "pandoc", row.names = F)
#datatable(res.mat)
```

Cost optimal is a buffer stock of `r res.mat$y[which.min(res.mat$total.cost)]` with associated total cost per month of `r round(res.mat$total.cost[which.min(res.mat$total.cost)],2)`  Euro.

Note that there are small differences due to rounding betweenthe values of column "total.cost" and the sum of the columns "cost.os" and "cost.us".

