---
title: "Solution Exercise 6 - Continuous review policy"
author: "Thomas Kirschstein"
subtitle: Inventory Management
output:
  pdf_document: default
  html_document:
    df_print: paged
    toc: yes
    toc_float: true
    number_sections: false
---
  
```{r, setup, echo=F, include=T, error=FALSE, warning=FALSE, message=FALSE}
library(DT)
library(kableExtra)
library(knitr)
library(forecast)
library(sarima)
library(gsl)

# sq inventory simulation
sq.inv.L <- function(d, s, q, l.ini, n, L){
  
  tmp.mat <- matrix(0, nrow= n+1+L, ncol= 5 )
  colnames(tmp.mat) <- c("demand","inv.level","inv.pos", "order", "delivery")
  
  tmp.mat[,"demand"] <- c(0,d,rep(0,L))
  tmp.mat[1,"Lagerbestand"] <- l.ini
  tmp.mat[1,"inv.pos"] <- l.ini
  
  for(i in 2:(n+1)){
    
    if(tmp.mat[i-1, "inv.pos"] <= s){
      tmp.mat[i, "order"] <- q
      tmp.mat[i+L, "delivery"] <- q
    }
    
    tmp.mat[i, "inv.pos"] <- tmp.mat[i-1, "inv.pos"] - tmp.mat[i, "demand"] + tmp.mat[i, "order"]
    tmp.mat[i, "inv.level"] <- tmp.mat[i-1, "inv.level"] - tmp.mat[i, "demand"] + tmp.mat[i, "delivery"]
  }
  return(tmp.mat[2:(n+1),]) 
}

# sq inventory optimization
sq.inv.opt <- function(pars, demand, wwbz, llini, obj = "cost", co = 0, cu = 0, B = 0, alpha.z = 95, beta.z = 99){
  
  sim.R <- sq.lager(d = demand, s = pars[1], q = pars[2], l.ini = llini, n = length(demand), L = wwbz)
  
  tmp <- sim.R[-c(1),]
  tmp.l <- tmp[,"inv.level"]
  tmp.id <- 1:length(tmp.l)
  tmp.id.sub <- tmp[,"delivery"] > 0
  tmp.id.sub.w <- which(tmp.id.sub) - 1
  tmp.id.sub.w <-tmp.id.sub.w[tmp.id.sub.w >0]
  
  # Alpha-SL
  sim.alpha <- (100*(1 - sum( tmp.l[tmp.id.sub.w] < 0)/length(tmp.id.sub.w)))
  sim.alpha.diff <- (alpha.z - sim.alpha)^2
  # Beta-SL
  sim.beta <- (100*(1 - sum(-tmp.l[ tmp.l < 0])/sum(tmp[,"demand"])))
  sim.beta.diff <- (beta.z - sim.beta)^2
  
  # cost
  sim.cost <- ( sum(tmp[,"order"] > 0)*B +  sum(tmp.l[ tmp.l > 0]) * co + sum(-tmp.l[ tmp.l < 0]) * cu)/length(tmp.l)
  
  return(ifelse(obj == "cost", sim.cost, ifelse(obj == "alpha", sim.alpha.diff, sim.beta.diff)) )
  
}

# Normal loss functions - direct
v.fun.direct.norm <- function(s, mu = 0, sigma = 1){
  integrate(function(y) (y-s)*dnorm(y, mean=mu, sd = sigma), lower = s, upper=Inf)$value
}

# deviation Beta vs. exp. loss 
v.diff.norm <- function(x, beta, mu = 0, sigma = 1, L,  c.or, c.sh){
  (v.fun.direct.norm(x[1], mu = L*mu, sigma = sqrt(L)*sigma) - (1-beta)*x[2])
}


v.diff.norm.q <- function(s, q, beta, mu = 0, sigma = 1, L,  c.or, c.sh){
  (v.fun.direct.norm(s, mu = L*mu, sigma = sqrt(L)*sigma) - (1-beta)*q)^2
}


################################################
# Optimization under beta service level constraint

sq.cost <- function(x, c.or, c.sh,mu, L, beta, sigma){
  c.sh *(x[2]/2 + x[1] - mu*L) + c.or*mu/x[2]
}


iter.sq <- function(mu = 100, sigma = 30, L = 8, c.or = 120, c.sh = 0.024, beta = .95){
  
  iter <- 1
  q.opt.old <- Inf
  lambda.opt <- 0
  ef <- 0

  repeat{
    q.opt <- sqrt(2*(mu*c.or+lambda.opt*ef) /(c.sh))
    if( abs(q.opt.old - q.opt) <= 1e-4 )  return(c(s.opt, q.opt))
    ef <- (1-beta)*q.opt
    s.opt <- optim(fn = v.diff.norm.q, par = mu*L, lower = 0, upper = mu*L*5, method="L-BFGS-B", mu = mu, sigma = sigma, q = q.opt, beta = beta, L = L)$par
    lambda.opt <- c.sh*q.opt/(1-pnorm(s.opt, mean = L*mu, sd = sqrt(L)*sigma ))
    res <- c(iter, q.opt, ef, s.opt, lambda.opt)
    names(res) <- c("iter","q","ef","s","lambda")
    print(res)
    q.opt.old <- q.opt
    iter <- iter + 1
    if(iter > 100) break
  }
  
}



v.fun.direct.gamma <- function(s, mu = 0, sigma = 1){
  integrate(function(y) (y-s)*dgamma(y, shape = mu^2/sigma^2, rate = mu/sigma^2), lower = s, upper=Inf)$value
}

v.fun.gamma <- function(s, mu = 0, sigma = 1){
  alpha <- mu^2/sigma^2
  beta <- mu/sigma^2
  alpha/beta * (1 - pgamma(s, shape = alpha+1, rate = beta)) - s*(1 - pgamma(s, shape = alpha, rate = beta))
}


iter.sq.mult <- function(mu = 100, sigma = 30, L = 8, c.or = 120, c.sh = 0.024, beta = .95, dist = "norm"){
  
  # beta vs. loss normal
  v.diff.norm.q <- function(s, q, beta, mu = 0, sigma = 1, L,  c.or, c.sh){
    (v.fun.direct.norm(s, mu = L*mu, sigma = sqrt(L)*sigma) - (1-beta)*q)^2
  }
  # beta vs. loss gamma
  v.diff.gamma.q <- function(s, q, beta, mu = 0, sigma = 1, L,  c.or, c.sh){
    (v.fun.gamma(s, mu = L*mu, sigma = sqrt(L)*sigma) - (1-beta)*q)^2
  }
  
  if(!(dist %in% c("gamma","norm"))) stop("Neither normal nor gamma distribution defined")
     
  iter <- 1
  q.opt.old <- Inf
  lambda.opt <- 0
  ef <- 0
  
  repeat{
    q.opt <- sqrt(2*(mu*c.or+lambda.opt*ef) /(c.sh))
    if( abs(q.opt.old - q.opt) <= 1e-4 )  return(c(s.opt, q.opt))
    ef <- (1-beta)*q.opt
    if(dist == "norm"){
      s.opt <- optim(fn = v.diff.norm.q, par = mu*L, lower = 0, upper = mu*L*5, method="L-BFGS-B", mu = mu, sigma = sigma, q = q.opt, beta = beta, L = L)$par
      lambda.opt <- c.sh*q.opt/(1-pnorm(s.opt, mean = L*mu, sd = sqrt(L)*sigma ))
    }
    if(dist == "gamma"){
      s.opt <- optim(fn = v.diff.gamma.q, par = mu*L, lower = 0, upper = mu*L*5, method="L-BFGS-B", mu = mu, sigma = sigma, q = q.opt, beta = beta, L = L)$par
      lambda.opt <- c.sh*q.opt/(1-pgamma(s.opt, shape = (L*mu)^2/L/sigma^2, rate = L*mu/L/sigma^2 ))
    }
      
    res <- c(iter, q.opt, ef, s.opt, lambda.opt)
    names(res) <- c("iter","q","ef","s","lambda")
    #print(res)
    q.opt.old <- q.opt
    iter <- iter + 1
    if(iter > 100) break
  }
  
}

```
	
# Continuous review with fixed lead time (I)

1. Calculate the cost-optimal order quantity $q$ and re-order level $s$ assuming an order lead time of 1 week and a minimum $\beta$ service level of 97.5%.

We use the algorithm presented in the lecture. In the first iteration we start with $q^0=\sqrt{\frac{2 \cdot 50 \cdot 80}{0.05}} = 400$, $l^0 = 0.025 \cdot 400 = 10$, $s = 40.45$, and $\lambda^0 = 22.64$ as follows:

```{r, exe1-1, echo=FALSE, message=TRUE, error=FALSE}

iter.sq(mu = 50, sigma = 8, c.sh = 0.05, c.or = 80, beta = .975, L = 1)

```

2. What happens with the re-order level $s$ if the order lead time reduces to 3 days? (no calculations required)

The reorder level decreases as the expected demand in the lead time decreases. However, $s$ decreases non-linearly as the standard deviation scales by $\sqrt{\frac{3}{5}} > \frac{3}{5}$.

# Continuous review with fixed lead time  (II)

A warehouse operator is observing the demand of a particular product. The operator estimates an expected daily demand of 10 and a standard deviation of 5. However, she is not sure which underlying demand distribution is appropriate; a normal distribution or a Gamma distribution are considered as likely candidates. The operator pursues a $(s,q)$ policy for the product which shows a reorder time of 1 day. 

Determine the re-order point $s$ for a normal and Gamma distribution of demand assuming that an $\alpha$ or a $\beta$ service level of 99% should by realized. The order quantity is set to $q=80$ units.

For the normal distribution, we have $s^{\alpha} = 10 + 2.3263 \cdot 5 = 21.63$ and $s^{\beta} = 10 + 0.63 \cdot 5 = 13.15$ whereby $L^{-1}(0.16)\approx 0.63$.

For the gamma distribution, we have $s^{\alpha}= 25.2542$ and $s^{\beta} \approx 19.7$.


# Continuous review with stochastic lead time  

Assume that the daily demand of cement on a construction side is Gamma distributed. The expected daily demand is estimated to be 8 tons and the standard deviation is supposed to be 6 tons. When new cement is ordered, the lead time varies. Based on historical records, with 20% probability the delivery arrives after 2 days. A delay of 3 days is observed in 50% of all cases and with 30% probability 4 days expire before the shipment arrives. 

1. Determine the distributional parameters for all three different lead time realizations.  

The distribution parameters for daily demand are $\alpha = \frac{\mu^2}{\sigma^2} = 1.778$ and $\beta =\frac{\mu}{\sigma^2}=0.222$. If the demand is i.i.d., it follows $Y^l \sim \Gamma\left(\sum_{i=1}^l \alpha_i , \beta\right)$

| $l$ 	|  2	| 3 	|  4	|
|:-:	|:-:	|:-:	|:-:	|
| $\sum_{i=1}^l \alpha_i$ 	|  $3.556$	|  $5.333$	| $7.111$ 	|

Thus, $Y^{l}\sim \Gamma(l\cdot 1.778,0.222)$.

2. Assume an $\alpha$ service level of 99% should be achieved. Which reorder level $s$ should be chosen?

The (probability weighted) demand distributions for all realizations of the lead time as well as joint demand distribution look like
```{r, exe3-2, echo=FALSE}
x.seq <- seq(0,100, length.out = 1000)
mu <- 8
sigma <- 6
alpha <- mu^2/sigma^2
beta <- mu/sigma^2
# lead time = 2
plot(x.seq, dgamma(x.seq , shape =  2*alpha, rate = beta )*0.2, type="l" , ylim=c(0,.035) , xlab="demand", ylab="probability", col=2)
# lead time = 3
lines(x.seq, dgamma(x.seq , shape =  3*alpha, rate = beta )*0.5, col=3 )
# lead time = 4
lines(x.seq, dgamma(x.seq , shape =  4*alpha, rate = beta )*0.3, col=4 )
# lead time = 4
lines(x.seq, dgamma(x.seq , shape =  4*alpha, rate = beta )*0.3 + dgamma(x.seq , shape =  3*alpha, rate = beta )*0.5 +dgamma(x.seq , shape =  2*alpha, rate = beta )*0.2 , col=1 )
legend("topright", legend=c("joint probability",paste("L=",2:4,sep="")), col = 1:4, lwd=2 , bty="n") 
```


Based on the tables, we estimate $s^\alpha$ for each realization of lead time:

| $l$ 	|  2	| 3 	|  4	|
|:-:	|:-:	|:-:	|:-:	|
| $P(L=l)$ 	|  $20%$	|  $50%$	| $30%$ 	|
| $s_l^\alpha$ 	|  $42.0$	|  $55.9$	| $66.6$ 	|

We conclude that $s^{\alpha}<s_4^\alpha =66.6$. Furthermore, we have $P(Y^{L=2}>66.6) \approx 0$, $P(Y^{L=3}>66.6) < 0.0015$ and $P(Y^{L=4}>66.6) = 0.01$, we can look for $x$ such that $P(Y^{L=4}>s) = 1-\frac{0.01}{0.3}$ such that $s^{\alpha} \approx 57.3$. 


